Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-04-13 Report/convergence_centralized.tex
final_antti_show_adhithya_all

In general, to ensure the uniqueness of the transmit precoders and the receive beamformers in all iterations, \textit{i.e}, even when the objective value is zero, the convex subproblems in \eqref{eqn-9} and \eqref{eqn-mse-2} can be regularized with a strongly convex function. Now, the objective function of the \me{\ith{i+1}} iteration is given as
\begin{equation}
\| \tilde{\mbf{v}} \|_q + c \, \| \mbf{x} - \mbf{x}^{(i)} \|^2
\end{equation}
where \me{\mbf{x}} is the vector containing all the optimization variables and \me{c} is a positive constant to make the objective strongly convex. The uniqueness of the minimizer is guaranteed by the strongly convex objective function \cite{scutari_1}. The vector \me{\mbf{x}^{(i)}} is the value of \me{\mbf{x}} obtained from the \me{\ith{i}} iteration.

@@

\subsection{Stationary Points of the Nonconvex Problem}
To show the limit point of the sequence of iterates \me{\{\iterate{\ma}{i}\}} is a stationary point, it must satisfy the \ac{KKT} conditions of the problem in \eqref{con}. As \me{i \rightarrow \infty}, the objective converges as
\iftoggle{single_column}{
\begin{equation} \label{con-opt}\allowdisplaybreaks
f(\iter{\mx}{\ast}{i},\iter{\my}{\ast}{i},\iter{\mz}{\ast|x}{i}) = f(\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i},\iter{\mz}{\ast|y}{i+1}) = 
f(\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i+1},\iter{\mz}{\ast|x}{i+1})
\end{equation}}{
\begin{multline} \label{con-opt}\allowdisplaybreaks
f(\iter{\mx}{\ast}{i},\iter{\my}{\ast}{i},\iter{\mz}{\ast|x}{i}) = f(\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i},\iter{\mz}{\ast|y}{i+1}) \\ = 
f(\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i+1},\iter{\mz}{\ast|x}{i+1})
\end{multline}}
where \me{\{\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i+1},\iter{\mz}{\ast|x}{i+1}\} \in \mc{F}^\ast} is the limit point of \me{\mc{A}} for a given initial point. As discussed in Appendix \ref{c-b}, the unique limit point of the sequence of iterates \me{\{\iterate{\ma}{i}\}} is a fixed point of \me{\mc{A}}. Using \cite[Th. 10]{lanckriet2009convergence} and \cite[Th. 2 and 11]{scutari_1}, we can show that the fixed points obtained by the algorithm \me{\mc{A}} under the mapping \me{f} is a stationary point of the nonconvex problem \eqref{con}. It can be verified by writing the \ac{KKT} expressions for the original problem in \eqref{con} and the relaxed one in \eqref{con-m} at the limit point \eqref{con-opt}, using the linear relaxation \eqref{con-relax} for the nonconvex constraint \cite{marks1978technical}.

>>>>

In general, to ensure the uniqueness of the transmit precoders and the receive beamformers in all iterations, \textit{i.e}, even when the objective value is zero, the convex subproblems in \eqref{eqn-9} and \eqref{eqn-mse-2} can be regularized with a strongly convex function. The objective function for the \me{\ith{i+1}} iteration can be regularized as
\begin{equation}
\| \tilde{\mbf{v}} \|_q + c \, \| \mbf{x} - \mbf{x}^{(i)} \|^2
\end{equation}
where \me{\mbf{x}} is the vector containing all the optimization variables and \me{c} is a positive constant to make the objective strongly convex. The uniqueness of the minimizer is guaranteed by the strongly convex objective function \cite{scutari_1}. The vector \me{\mbf{x}^{(i)}} is the value of \me{\mbf{x}} obtained from the \me{\ith{i}} iteration.

@@

\subsection{Stationary Points of the Nonconvex Problem}
To show the limit point of the sequence of iterates \me{\{\iterate{\ma}{i}\}} is a stationary point, it must satisfy the \ac{KKT} conditions of the problem in \eqref{con}. As \me{i \rightarrow \infty}, the objective converges as
\iftoggle{single_column}{
\begin{equation} \label{con-opt}\allowdisplaybreaks
f(\iter{\mx}{\ast}{i},\iter{\my}{\ast}{i},\iter{\mz}{\ast|x}{i}) = f(\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i},\iter{\mz}{\ast|y}{i+1}) = 
f(\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i+1},\iter{\mz}{\ast|x}{i+1})
\end{equation}}{
\begin{multline} \label{con-opt}\allowdisplaybreaks
f(\iter{\mx}{\ast}{i},\iter{\my}{\ast}{i},\iter{\mz}{\ast|x}{i}) = f(\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i},\iter{\mz}{\ast|y}{i+1}) \\ = 
f(\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i+1},\iter{\mz}{\ast|x}{i+1})
\end{multline}}
where \me{\{\iter{\mx}{\ast}{i+1},\iter{\my}{\ast}{i+1},\iter{\mz}{\ast|x}{i+1}\} \in \mc{F}^\ast} is the limit point of \me{\mc{A}} for a given initial point. As discussed in Appendix \ref{c-b}, the unique limit point of the sequence of iterates \me{\{\iterate{\ma}{i}\}} is a fixed point of \me{\mc{A}}. Using \cite[Th. 10]{lanckriet2009convergence} and \cite[Th. 2 and 11]{scutari_1}, we can show that the fixed points obtained by the algorithm \me{\mc{A}} under the mapping \me{f} is a stationary point of the nonconvex problem \eqref{con}. It can be verified by writing the \ac{KKT} expressions for the original problem in \eqref{con} and the relaxed one in \eqref{con-m} at the limit point \eqref{con-opt}, using the relation in \eqref{con-relax} between the nonconvex constraint and the corresponding convex approximation \cite{marks1978technical}.

### 2015-04-13 Report/introduction.tex
final_antti_show_adhithya_all

The relation between the achievable sum rate and the \ac{MSE} of the received symbol by using fixed \ac{MMSE} receivers can be used to solve the \ac{WSRM} problem \cite{mse_duality}. In \cite{christensen2008weighted,wmmse_shi}, the \ac{WSRM} problem is reformulated via \ac{MSE}, casting the problem as a convex one for fixed linearization coefficients. In this way, the original problem is expressed in terms of the \ac{MSE} weight, precoders, and decoders. Then the problem is solved using an alternating optimization method, i.e., finding a subset of variables while the other variables are fixed. The \ac{MSE} reformulation for the \ac{WSRM} problem has also been studied in \cite{hong2012decomposition} by using the \ac{SCA} to solve the problem in an iterative manner. Additionally the rate constraints based on the \ac{QoS} requirements are included in the \ac{WSRM} problem and solved via the \ac{MSE} reformulation in \cite{kaleva2013primal,kaleva2013decentralized}.

@@

By adjusting the weights in the \ac{WSRM} objective, we can find an arbitrary rate-tuple in the rate region that maximizes suitable objective measures. For example, if the weight of each user is set to be inversely proportional to its average data rate, the corresponding problem guarantees fairness on the average among the users. To reduce the number of backlogged packets, we can assign weights based on the current queue size of the users. Specifically, the queue states can be incorporated in the \ac{WSRM} objective $\sum_k w_k R_k$ by replacing the weight $w_k$ with the corresponding queue state $Q_k$ or its function, which is the outcome of minimizing the Lyapunov drift between the current and the future queue states \cite{tassiulas,neely2010stochastic}, where \me{R_k} denotes the achievable data rate of user \me{k}. In the backpressure algorithm, the differential queues between the source and the receiver nodes are used to scale the transmission rate \cite{georgiadis2006resource}.

Earlier studies on the queue minimization problem were summarized in the survey papers \cite{berry2004cross,layering_as_opt}. In particular, the problem of power allocation to minimize the number of backlogged packets was considered in \cite{qps_cioffi} using geometric programming. Since the problem addressed in \cite{qps_cioffi} assumed single antenna transmitters and receivers, the queue minimizing problem reduces to the optimal power allocation problem. In the context of wireless networks, the backpressure algorithm mentioned above was extended in \cite{weeraddana2011resource} by formulating the corresponding user queues as the weights in the \ac{WSRM} problem. Recently, the precoder design for the video transmission over \ac{MIMO} system is considered in \cite{video_queues}. In this design, the \ac{MU}-\ac{MIMO} precoders are designed by the \ac{MSE} reformulation as in \cite{christensen2008weighted} with the higher layer performance objective such as playback interruptions and buffer overflow probabilities.

\subsubsection*{Main Contributions}
\review{In this paper, we design precoders jointly over space-frequency resources to reduce the number of backlogged packets waiting at each \acp{BS}. The proposed formulation also limits the allocations beyond the number of backlogged packets without explicit rate constraints. Initially, we propose centralized \ac{JSFRA} formulation, which is solved by two iterative algorithms based on the combination of \ac{SCA} and \ac{AO} due to the nonconvex nature of the problem. The proposed algorithms solve a sequence of convex problems obtained by fixing a subset of optimization variables or by approximating the nonconvex constraints by the convex ones. The first approach is performed by directly relaxing the \ac{SINR} expression, while in the second technique the equivalence between the \ac{MSE} and the \ac{SINR} is exploited.} We then discuss the distributed implementation of the \ac{JSFRA} methods using primal decomposition and the \ac{ADMM}. Finally, we also propose a more practical iterative precoder design by directly solving the \ac{KKT} system of equations for the \ac{MSE} reformulation that is numerically shown to require minimal information exchange for each update. \review{Note that the joint space-frequency channel matrix can be formed by stacking the channel of each sub-channel in a block-diagonal form for all users}.

>>>>

The relation between the achievable sum rate and the \ac{MSE} of the received symbol by using fixed \ac{MMSE} receivers can be used to solve the \ac{WSRM} problem \cite{mse_duality}. In \cite{christensen2008weighted,wmmse_shi}, the \ac{WSRM} problem is reformulated via \ac{MSE}, casting the problem as a convex one for fixed linearization coefficients. In this way, the original problem is expressed in terms of the \ac{MSE} weight, precoders, and decoders. Then the problem is solved using an alternating optimization method, i.e., finding a subset of variables while the other variables are fixed. The \ac{MSE} reformulation for the \ac{WSRM} problem has also been studied in \cite{hong2012decomposition} by using \ac{SCA} to solve the problem in an iterative manner. Moreover, distributed precoder design with \ac{QoS} requirements as additional rate constraints are studied for the \ac{MSE} reformulated \ac{WSRM} problem in \cite{kaleva2013primal,kaleva2013decentralized}.

@@

By adjusting the weights in the \ac{WSRM} objective, we can find an arbitrary rate-tuple in the rate region that maximizes suitable objective measures. For example, if the weight of each user is set to be inversely proportional to its average data rate, the corresponding problem guarantees fairness on the average among the users. To reduce the number of backlogged packets, we can assign weights based on the current queue size of the users. Specifically, the queue states can be incorporated in the \ac{WSRM} objective $\sum_k w_k R_k$ by replacing the weight $w_k$ with the corresponding queue state $Q_k$ or its function, which is the outcome of minimizing the Lyapunov drift between the current and the future queue states \cite{tassiulas,neely2010stochastic}, where \me{R_k} denotes the achievable data rate of user \me{k}. In the \textit{backpressure algorithm}, the differential queues between the source and the receiver nodes are used to scale the transmission rate \cite{georgiadis2006resource}.

Earlier studies on the queue minimization problem are summarized in the survey papers \cite{berry2004cross,layering_as_opt}. In particular, the problem of power allocation to minimize the number of backlogged packets is considered in \cite{qps_cioffi} using geometric programming. Since the problem addressed in \cite{qps_cioffi} assumed single antenna transmitters and receivers, the queue minimizing problem reduces to the optimal power allocation problem. In the context of wireless networks, the \textit{backpressure algorithm} mentioned above is extended in \cite{weeraddana2011resource} by formulating the corresponding user queues as the weights in the \ac{WSRM} problem. Recently, the precoder design for the video transmission over \ac{MIMO} system is considered in \cite{video_queues}. In this design, the \ac{MU}-\ac{MIMO} precoders are designed by the \ac{MSE} reformulation as in \cite{christensen2008weighted} with the higher layer performance objective such as playback interruptions and buffer overflow probabilities.

\subsubsection*{Main Contributions}
In this paper, we design precoders jointly over space-frequency resources to reduce the number of backlogged packets waiting at each \acp{BS}. The proposed formulation also limits the allocations beyond the number of backlogged packets without explicit rate constraints. \review{Initially, we propose a centralized \ac{JSFRA} formulation, which is solved by two iterative algorithms based on the combination of \ac{SCA} and \ac{AO} due to the nonconvex nature of the problem. The proposed algorithms solve a sequence of convex problems obtained by fixing a subset of optimization variables or by approximating the nonconvex constraints by the convex ones. The first approach is performed by directly relaxing the \ac{SINR} expression, while in the second method, the equivalence between the \ac{MSE} and the \ac{SINR} is exploited.} We then discuss the distributed implementation of the \ac{JSFRA} methods using primal decomposition and the \ac{ADMM}. Finally, we also propose a more practical iterative precoder design by directly solving the \ac{KKT} system of equations for the \ac{MSE} reformulation that is numerically shown to require minimal information exchange for each update. \review{Note that the joint space-frequency channel matrix can be formed by stacking the channel of each sub-channel in a block-diagonal form for all users}.

### 2015-04-13 Report/system_model.tex
final_antti_show_adhithya_all

We adopt linear transmit beamforming technique at \acp{BS}. Specifically, the data symbols \me{d_{l,k,n}} for  user \me{k} on the \me{l^\mathrm{th}} spatial stream over sub-channel \me{n} is multiplied with beamformer \me{\mvec{m}{l,k,n} \in \mathbb{C}^{N_T \times 1}} before being transmitted. In order to detect  multiple spatial streams at the user terminal, receive beamforming vector \me{\mvec{w}{l,k,n}} is employed for each user. Consequently, the received data estimate corresponding to the \me{l^\mathrm{th}} spatial stream over sub-channel \me{n} at user $k$ is given by
\iftoggle{single_column}{
\begin{equation}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n}
\end{equation}
}{
\begin{multline}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n} \\ 
+ \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n}
\end{multline}
}
where \me{\mvec{H}{b,k,n} \in \mathbb{C}^{N_R \times N_T}} is the channel between \ac{BS} \me{b} and  user \me{k} on sub-channel \me{n}, and \me{\mvec{n}{k,n}} \me{\sim \mc{CN}(0,N_0)} is the additive noise vector for user \me{k} on the \me{n^\mathrm{th}} sub-channel and \me{l^\mathrm{th}} spatial stream. In \eqref{eqn-1}, \me{L = \text{rank}(\mvec{H}{b,k,n}) = \min(N_T,N_R)} is the maximum number of spatial streams.\footnote{It can be easily extended for user specific streams \me{L_k} instead of using common \me{L} streams for all users. \me{L} streams are initialized but after solving the problem, only \me{L_{k,n} \leq L} non-zero data streams are transmitted.} Assuming independent detection of data streams, we can write the signal-to-interference-plus-noise ratio (SINR) as
\begin{equation}\label{eq:SINR}
\gamma_{l,k,n} = \dfrac{\left |\mvec{w}{l,k,n}^\herm \, \mvec{H}{b_k,k,n} \, \mvec{m}{l,k,n} \right |^2}{\enoise + \sum_{(j,i) \neq (l,k)} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2}
\end{equation}
where \me{\enoise = N_0 \, \trace (\mvec{w}{l,k,n} \mbf{w}_{l,k,n}^\herm)} denotes the equivalent noise variance. \review{To reduce the overhead involved in feeding back the user channels, we consider a \ac{TDD} system in which \acp{BS} can estimate the downlink channels from the uplink ones using the so-called channel reciprocity.}

>>>>

We adopt linear transmit beamforming technique at \acp{BS}. Specifically, the data symbols \me{d_{l,k,n}} for  user \me{k} on the \me{l^\mathrm{th}} spatial stream over sub-channel \me{n} is multiplied with beamformer \me{\mvec{m}{l,k,n} \in \mathbb{C}^{N_T \times 1}} before being transmitted. In order to detect  multiple spatial streams at the user terminal, receive beamforming vector \me{\mvec{w}{l,k,n}} is employed for each user. Consequently, the received data estimate corresponding to the \me{l^\mathrm{th}} spatial stream over sub-channel \me{n} at user $k$ is given by
\iftoggle{single_column}{
\begin{equation}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n}
\end{equation}
}{
\begin{multline}\label{eqn-1}
\hat{d}_{l,k,n} = \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \,\mvec{m}{l,k,n} d_{l,k,n} + \mvec{w}{l,k,n}^\herm \mvec{n}{k,n} \\ 
+ \mvec{w}{l,k,n}^\herm \sum_{i \in \mc{U} \backslash \set{k}} \mvec{H}{b_i,k,n} \sum_{j = 1}^L \mvec{m}{j,i,n}d_{j,i,n}
\end{multline}
}
where \me{\mvec{H}{b,k,n} \in \mathbb{C}^{N_R \times N_T}} is the channel between \ac{BS} \me{b} and  user \me{k} on sub-channel \me{n}, and \me{\mvec{n}{k,n}} \me{\sim \mc{CN}(0,N_0)} is the additive noise vector for user \me{k} on the \me{n^\mathrm{th}} sub-channel and \me{l^\mathrm{th}} spatial stream. In \eqref{eqn-1}, \me{L = \text{rank}(\mvec{H}{b,k,n}) = \min(N_T,N_R)} is the maximum number of spatial streams.\footnote{It can be easily extended for user specific streams \me{L_k} instead of using common \me{L} streams for all users. \me{L} streams are initialized but after solving the problem, only \me{L_{k,n} \leq L} non-zero data streams are transmitted.} Assuming independent detection of data streams, we can write the signal-to-interference-plus-noise ratio (SINR) as
\begin{equation}\label{eq:SINR}
\gamma_{l,k,n} = \dfrac{\left |\mvec{w}{l,k,n}^\herm \, \mvec{H}{b_k,k,n} \, \mvec{m}{l,k,n} \right |^2}{\enoise + \sum_{(j,i) \neq (l,k)} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2}
\end{equation}
where \me{\enoise = N_0 \, \trace (\mvec{w}{l,k,n} \mbf{w}_{l,k,n}^\herm)} denotes the equivalent noise variance. \review{To reduce the overhead involved in feeding back the user channels, we consider a \ac{TDD} system in which \acp{BS} can estimate the downlink channels from the uplink pilots using channel reciprocity.}
